---
title: "Age-Based Index (ABI) and Blim Methods in FLRebuild"
author: "Package Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Age-Based Index (ABI) and Blim Methods in FLRebuild}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

This vignette demonstrates the use of the Age-Based Index (ABI) and Blim methods in the `FLRebuild` package. These methods provide tools for analyzing the age structure of fish stocks and for calculating biomass limit reference points, which are important for fisheries management and stock assessment.

## Setup

```{r setup}
library(FLRebuild)
library(FLCore)
library(FLBRP)
```

## Example Data

We will use the example datasets `ple4` (an `FLStock` object) and `ple4brp` (an `FLBRP` object) provided by the `FLCore` and `FLBRP` packages.

```{r data}
data(ple4, package = "FLCore")
data(ple4brp, package = "FLBRP")
```

## ABI Methods

The ABI (Age-Based Index) methods allow you to:
- Calculate the reference age at which a given proportion of the cumulative stock numbers is reached (`abiAge`)
- Calculate the proportion of stock numbers above the reference age at MSY (`abiMsy`)
- Calculate the observed proportion above the reference age for an `FLStock` object, relative to the MSY reference (`abi`)

### Calculate Reference Age (`abiAge`)

```{r abiAge}
ref_age <- abiAge(ple4brp)
ref_age
```

### Calculate Proportion Above Reference Age at MSY (`abiMsy`)

```{r abiMsy}
p_above_ref <- abiMsy(ple4brp)
p_above_ref
```

### Calculate Observed Proportion Above Reference Age (`abi`)

```{r abi}
abi_obs <- abi(ple4, ple4brp)
abi_obs
```

### Visualizing Age Structure

You can visualize the age structure and the reference age threshold:

```{r plot-age-structure}
# Extract stock numbers at the most recent year
stk_n <- stock.n(ple4)[, dim(ple4)[2], drop=TRUE]
ages_vec <- as.numeric(dimnames(stk_n)$age)

barplot(as.numeric(stk_n), names.arg = ages_vec,
        main = "Stock Numbers by Age (Most Recent Year)",
        xlab = "Age", ylab = "Stock Numbers")
abline(v = as.numeric(ref_age), col = "red", lwd = 2, lty = 2)
legend("topright", legend = c("Reference Age"), col = "red", lty = 2, lwd = 2)
```

## Blim Method

The `blim` method calculates biomass limit reference points (Blim) for an `FLBRP` object based on a ratio of virgin recruitment. This is useful for defining biological reference points in management strategy evaluation.

### Calculate Blim Reference Points

```{r blim}
blim_ref <- blim(ple4brp, ratio = 0.3)
blim_ref
```

## Interpretation

- **ABI methods** help you understand the age structure of the stock and how much of the population is above a biologically meaningful age threshold.
- **Blim** provides a reference point for the minimum safe biomass, below which the stock is considered to be at risk.

## References

- Kell, L. T., Mosqueira, I., et al. (2005). Evaluation of multiple management objectives for mixed fisheries. ICES Journal of Marine Science, 62(6), 1236-1252.
- ICES (2018). ICES Advice basis. https://www.ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/Advice_basis_2018.pdf
- FLR Project: https://flr-project.org/ 