---
title: "Introduction to FLRebuild"
author: "Package Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to FLRebuild}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

FLRebuild is an R package designed for analyzing fisheries stock rebuilding trajectories and calculating rebuilding times. It provides methods for projecting stock rebuilding from different initial depletion levels using FLR (Fisheries Library for R) objects.

## Installation

```{r eval=FALSE}
# Install from GitHub (if available)
# devtools::install_github("username/FLRebuild")

# Or install from local source
# install.packages("path/to/FLRebuild", repos = NULL, type = "source")
```

## Basic Usage

### Loading the Package

```{r}
library(FLRebuild)
```

### Working with FLBRP Objects

The package provides methods for working with FLBRP (Fisheries Library for R - Biological Reference Points) objects:

```{r eval=FALSE}
# Create a life history equilibrium object
library(FLife)
eq <- lhEql(lhPar(FLPar(linf=250, s=0.9)))

# Run rebuilding analysis
stk <- rebuild(eq, nInitial = 50)

# Calculate rebuilding times
rebuild_times <- rebuildTime(stk)
```

### Working with biodyn Objects

For biodyn objects, you can calculate rebuilding trajectories:

```{r eval=FALSE}
# Create a biodyn object
library(biodyn)
bd <- biodyn(FLPar(r=0.5, k=1000, p=1))

# Calculate rebuilding time
rebuild_data <- rebuildTime(bd, target = 1000)
```

### Utility Functions

The package also includes utility functions for calculating Tmax (maximum rebuilding time) using different methods as described in NS1 guidelines:

```{r}
# Calculate Tmax using generation time method
tmax_gen <- calculateTmax(object = NULL, method = "generation", tmin = 10)

# Calculate Tmax using multiply method
tmax_mult <- calculateTmax(object = NULL, method = "multiply", tmin = 10)

# Calculate recovery time based on growth rate
recovery_time <- calculateRecoveryTime(initial_biomass = 0.5, 
                                      target_biomass = 1.0, 
                                      growth_rate = 0.1)
```

## Key Features

1. **Multiple Object Types**: Works with FLBRP, biodyn, and FLStock objects
2. **Flexible Rebuilding Analysis**: Projects trajectories from various initial depletion levels
3. **Tmax Calculation**: Implements different methods for calculating maximum rebuilding time
4. **Visualization**: Includes plotting functions for rebuilding trajectories
5. **NS1 Guidelines Compliance**: Follows National Standard 1 guidelines for rebuilding analysis

## Methods for Tmax Calculation

The package implements three methods for calculating Tmax as described in NS1 guidelines:

1. **Generation Method**: Tmin + one generation time
2. **Fishing Method**: Time to rebuild at 75% of Maximum Fishing Mortality Threshold (MFMT)
3. **Multiply Method**: Tmin multiplied by 2

## Examples

### Basic Rebuilding Analysis

```{r eval=FALSE}
# Load required packages
library(FLRebuild)
library(FLife)

# Create equilibrium object
eq <- lhEql(lhPar(FLPar(linf=250, s=0.9)))

# Run rebuilding analysis
stk <- rebuild(eq, 
               targetF = 0,           # No fishing during rebuild
               targetSSB = 1000,      # Target SSB
               nInitial = 100,        # Number of initial conditions
               burnin = 20)           # Burn-in period

# Calculate rebuilding times
rebuild_times <- rebuildTime(stk, nx = 101)

# Plot results
plotRebuildTrajectories(as.data.frame(ssb(stk), drop = TRUE))
```

### Advanced Analysis

```{r eval=FALSE}
# Analyze multiple scenarios
scenarios <- list(
  no_fishing = list(targetF = 0),
  low_fishing = list(targetF = 0.1),
  moderate_fishing = list(targetF = 0.2)
)

results <- lapply(scenarios, function(scen) {
  rebuild(eq, targetF = scen$targetF, nInitial = 50)
})

# Compare rebuilding times
rebuild_times <- lapply(results, rebuildTime)
```

## Contributing

Contributions are welcome! Please feel free to submit issues, feature requests, or pull requests.

## License

This package is licensed under the GPL-3 license. 