---
title: "Parameterising a Ricker Stock-Recruitment Relationship by the Location of the Peak"
author: "Laurie Kell"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

## Overview

The standard Ricker stock-recruitment relationship can be written as

$$
R(S) = \alpha S \exp(-\beta S)
$$

where:

-   $R(S)$ is recruitment at spawning-stock biomass $S$;
-   $\alpha$ controls productivity at low stock size;
-   $\beta$ controls the strength of density dependence.

This note shows how to parameterise the Ricker relationship so that the peak in recruitment occurs at a specified spawning biomass, $S_{\text{peak}}$.

## Location of the peak

To find the spawning biomass at which recruitment is maximised, differentiate $R(S)$ with respect to $S$:

$$
\frac{dR}{dS}
=
\alpha \exp(-\beta S)(1 - \beta S).
$$

The maximum occurs when

$$
1 - \beta S = 0.
$$

Therefore, the peak occurs at

$$
S_{\text{peak}} = \frac{1}{\beta}.
$$

So, if we want the peak of the Ricker curve to occur at a specified spawning biomass $S_{\text{peak}}$, we set

$$
\beta = \frac{1}{S_{\text{peak}}}.
$$

The Ricker relationship can then be written as

$$
R(S) =
\alpha S
\exp\left(-\frac{S}{S_{\text{peak}}}\right).
$$

## Parameterisation using maximum recruitment

If we want to parameterise the curve in terms of the maximum recruitment $R_{\max}$, rather than $\alpha$, note that at the peak

$$
R_{\max}
=
R(S_{\text{peak}})
=
\alpha S_{\text{peak}} \exp(-1).
$$

Rearranging gives

$$
\alpha =
\frac{e R_{\max}}{S_{\text{peak}}}.
$$

Substituting this into the Ricker equation gives the peak-parameterised form

$$
R(S) =
R_{\max}
\left(\frac{S}{S_{\text{peak}}}\right)
\exp\left(1 - \frac{S}{S_{\text{peak}}}\right).
$$

This form has two intuitive parameters:

-   $S_{\text{peak}}$, the spawning biomass at which recruitment is maximised;
-   $R_{\max}$, the maximum recruitment.

## R implementation

The peak-parameterised Ricker curve can be implemented as follows:

```{r ricker-peak-function}
ricker_peak <- function(S, S_peak, R_max) {
  R_max * (S / S_peak) * exp(1 - S / S_peak)
}
```

Alternatively, if retaining the original $\alpha, \beta$ notation, set

$$
\beta = \frac{1}{S_{\text{peak}}}
$$

and use:

```{r ricker-alpha-beta-function}
ricker_alpha_beta <- function(S, alpha, S_peak) {
  beta <- 1 / S_peak
  alpha * S * exp(-beta * S)
}
```

Thus, fixing the location of the peak is equivalent to fixing $\beta$, while $\alpha$ determines the vertical scale of the curve.

## Example

Suppose recruitment is to peak at $S_{\text{peak}} = 100$, with $R_{\max} = 1000$.

```{r example-parameters}
S_peak <- 100
R_max  <- 1000

S <- seq(0, 300, length.out = 200)
R <- ricker_peak(S, S_peak = S_peak, R_max = R_max)
```

The curve can be plotted as follows:

```{r example-plot, fig.width=7, fig.height=5}
plot(
  S, R,
  type = "l",
  lwd = 2,
  xlab = "Spawning-stock biomass, S",
  ylab = "Recruitment, R(S)",
  main = "Ricker stock-recruitment relationship"
)

abline(v = S_peak, lty = 2, col = "red")
abline(h = R_max,  lty = 2, col = "blue")

points(S_peak, R_max, pch = 16, col = "red")

legend(
  "topright",
  legend = c(expression(S[peak]), expression(R[max])),
  lty = 2,
  col = c("red", "blue"),
  bty = "n"
)
```

## Relationship between $\alpha$, $S_{\text{peak}}$, and $R_{\max}$

Given $S_{\text{peak}}$ and $R_{\max}$, the equivalent value of $\alpha$ is

$$
\alpha =
\frac{e R_{\max}}{S_{\text{peak}}}.
$$

In R:

```{r alpha-from-peak}
alpha_from_peak <- function(S_peak, R_max) {
  exp(1) * R_max / S_peak
}

alpha <- alpha_from_peak(S_peak, R_max)
alpha
```

The corresponding value of $\beta$ is

$$
\beta = \frac{1}{S_{\text{peak}}}.
$$

```{r beta-from-peak}
beta <- 1 / S_peak
beta
```

Using these values in the original Ricker form gives the same curve:

```{r compare-forms}
R_ab <- alpha * S * exp(-beta * S)

all.equal(R, R_ab)
```

## Summary

For a Ricker stock-recruitment relationship

$$
R(S) = \alpha S \exp(-\beta S),
$$

the peak occurs at

$$
S_{\text{peak}} = \frac{1}{\beta}.
$$

Therefore, to place the peak at a chosen spawning biomass $S_{\text{peak}}$, set

$$
\beta = \frac{1}{S_{\text{peak}}}.
$$

If the desired maximum recruitment is $R_{\max}$, then

$$
\alpha =
\frac{e R_{\max}}{S_{\text{peak}}},
$$

giving the equivalent peak-parameterised form

$$
R(S) =
R_{\max}
\left(\frac{S}{S_{\text{peak}}}\right)
\exp\left(1 - \frac{S}{S_{\text{peak}}}\right).
$$
